\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-01-25 results.tex
updates

\begin{figure}[htpb]
\centering
\includegraphics[width=0.5\textwidth]{figs/inl.jpg}
\caption{\small An Intermediate Nepheloid Layer (INL) detected by an
  AUV mission in the Monterey Bay in August 2005, showing the location
  of the AUV transect in the bay (left) and optical backscatter image
  at the 420 Nm frequency showing the INLs patchy nature
  (right). \small{Courtesy: John Ryan/MBARI}}
\label{fig:inl}
\end{figure}

@@

To augment our understanding of such microbial communities, the Dorado
with \rx has been utilized in a range of activities in a
quasi-lagrangian mode. As the water mass is being tracked with a
drifting instrument, the vehicle is expected to move around it. While
a number of different patterns have been tested in simulation, the
majority of at-sea deployments have used a square pattern ($3X3$
km\textsuperscript{2} or a $1X1$ km\textsuperscript{2})\footnote{Often
  post-hoc reconstruction of the data set for visualization drives how
  much adaptation the vehicle can be allowed to undertake.}. Plan
synthesis occurs frequently, but within the context of a drifting
point generated by a drifter. Fig. \ref{fig:dta-setup} shows the
general setup of our experiments which can be tailored with the use of
various computational methods within the \texttt{DTA} app which is
contained within the ODSS (see more below) \cite{das11}. GPS data from
the free-floating drifter is sent via satellite and processed onshore
to determine a projection (speed and bearing) of the drifter. This
data is sent to the AUV; at the completion of the current survey, the
vehicle receives this data as a high-level goal which is decomposed
and synthesized as a survey plan. Currently, the vehicle completes the
survey without any further dynamic updates from the drifter. In order
to minimize the range of errors that the survey needs to encapsulate
(data latency or surface or sub-surface currents for those drifters
which have a drogue) the estimation is a projection in time and space
of the drifters centroid sent with this goal from the shore-side app.

\begin{figure}[htpb]
\centering
\includegraphics[width=0.35\textwidth]{figs/june11-drifter-follow-180-181.pdf}
\caption{\small{A June 2011 drifter cruise shown in the Earth
    frame. Deviations in the AUV transect are based on the frame of
    reference in addition to positional errors as shown in
    \cite{das11b}.}}
\label{fig:drifter-errors}
\end{figure}

Typically this has meant that the vehicle computes a set of waypoints
for the survey plan, moves to the start waypoint and then executes the
survey even as the centroid is moving. Positional error of the vehicle
and uncertainty in the speed and bearing of the vehicle\footnote{In
  our September 2010 experiment as the drifter moved over the Davidson
  Seamount \cite{clague10} deviation in the California current
  resulted in visible directional change that can be seen in
  Fig. \ref{fig:sep10-transects} at the end of the mission.} exist;
however as \cite{das11b} clearly articulates, the observations of the
survey to retain the environmental context require a change in
reference frames so that the patch-center tagged by a GPS-tracked
drifter remains within the survey perimeter at all times.

@@

\begin{figure*}
\centering 
\subfloat[]{\label{fig:sep11-temp}\includegraphics[width=0.5\textwidth]{figs/sep11-ww-temp.png}} 
\subfloat[]{\label{fig:sep11-sal}\includegraphics[width=0.5\textwidth]{figs/sep11-ww-salinity.png}}\\
\subfloat[]{\label{fig:sep11-flh}\includegraphics[width=0.5\textwidth]{figs/sep11-ww-flh.png}} 
\subfloat[]{\label{fig:sep11-n2}\includegraphics[width=0.5\textwidth]{figs/sep11-ww-n2.png}} 
\caption{\small }
  \label{fig:sept11-wirewalker}
\end{figure*}

>>>>

\begin{figure}[htpb]
\centering
\includegraphics[width=0.5\textwidth]{figs/inl.jpg}
\caption{\small An Intermediate Nepheloid Layer (INL) detected by an
  AUV mission in the Monterey Bay in August 2005, showing the location
  of the AUV transect in the bay (left) and optical backscatter image
  at the 420 Nm frequency showing the INLs patchy nature
  (right). \emph{Image courtesy: John Ryan, MBARI}}
\label{fig:inl}
\end{figure}

@@

To augment our understanding of such microbial communities, the Dorado
with \rx has been utilized in a range of activities in a
quasi-lagrangian mode. As the water mass is being tracked with a
drifting instrument, the vehicle is expected to move around it. While
a number of different patterns have been tested in simulation, the
majority of at-sea deployments have used a square pattern ($3X3$
km\textsuperscript{2} or a $1X1$ km\textsuperscript{2})\footnote{Often
  post-hoc reconstruction of the data set for visualization drives how
  much adaptation the vehicle can be allowed to undertake.}. Plan
synthesis occurs frequently, but within the context of a drifting
point generated by a drifter. Fig. \ref{fig:dta-setup} shows the
general setup of our experiments which can be tailored with the use of
various computational methods within the \texttt{DTA} app which is
contained within the Oceanographic Decision Support System (ODSS; see
below) \cite{das11}. GPS data from the free-floating drifter is sent
via satellite and processed onshore to determine a projection (speed
and bearing) of the drifter. This data is sent to the AUV; at the
completion of the current survey, the vehicle receives this data as a
high-level goal which is decomposed and synthesized as a survey
plan. Currently, the vehicle completes the survey without any further
dynamic updates from the drifter. In order to minimize the range of
errors that the survey needs to encapsulate (data latency or surface
or sub-surface currents for those drifters which have a drogue) the
estimation is a projection in time and space of the drifters centroid
sent with this goal from the shore-side app.

\begin{figure}[htpb]
\centering
\includegraphics[width=0.35\textwidth]{figs/june11-drifter-follow-180-181.pdf}
\caption{\small{A June 2011 drifter cruise shown in the Earth
    frame. Deviations in the AUV transect are based on the frame of
    reference in addition to positional errors \cite{das11b}.}}
\label{fig:drifter-errors}
\end{figure}

Typically this has meant that the vehicle computes a set of waypoints
for the survey plan, moves to the start waypoint and then executes the
survey even as the centroid is moving. Positional error of the vehicle
and uncertainty in the speed and bearing of the vehicle\footnote{In
  our September 2010 experiment as the drifter moved over the Davidson
  Seamount \cite{clague10} deviation in the California current
  resulted in visible directional change that can be seen in
  Fig. \ref{fig:sep10-transects} at the end of the mission.} exist;
however, the observations of the survey to retain the environmental
context require a change in reference frames so that the patch-center
tagged by a GPS-tracked drifter remains within the survey perimeter at
all times \cite{das11b}.

@@

These theoretic results were validated in a five-day long off-shore
deployment carried out in September 2010, 160 kms off the California
coast (Fig.~\ref{fig:sep10-ship}). A specialized drifter with a
genomic sensor \cite{scholin09} hanging $20$ m below, performing
in-situ identification of micro-organisms was the target. The
experiment was supported by crews on two support vessels, the R/V
\emph{Western Flyer} and the R/V \emph{Zephyr}. The \emph{Flyer}
visited the drifter every four hours to carry out a series of
ship-based sampling experiments and lab analysis on water samples to
ground-truth the drifter sensor data. The \emph{Zephyr} was meanwhile
focused on Lagrangian observation studies using the Dorado with the
goal to monitor the nutrient budget at the perimeter of a $1$km $X
1$km water patch around the advecting drifter while the AUV performed
a transformed box pattern around the drifter. A number of logistical
issues were kept in mind while designing and executing the
experiment. Each iteration began with the latest drifter update
(position and velocity) received from the drifter through a satellite
link which was transmitted to the vehicle for in-situ adaptation. \rx
used this input to compute the $5$ waypoints necessary for an
iteration of the box pattern, with the AUV traveling at constant
velocity in the earth frame. Waypoints were computed once at the
beginning of the survey with the AUV surfacing once for every survey,
with each survey lasting $\sim1$ to $1.5$ hours. A total of 60 surveys
were attempted over the course of the $5$
days. Fig. \ref{fig:sep10-transects} shows the overall transects over
the course of the mission, with gaps in between days when the AUV was
being recharged. Each deployment lasted $\sim12$ hours. The black dots
in the figure show the beginning of each iteration when drifter
updates were received by the AUV.

\begin{figure}[htpb]
\centering
\includegraphics[width=0.7\textwidth]{figs/oct10-ballet.pdf}
\caption{\small An October 2010 robot ``ballet'' where a smaller
  slower AUV tracked a patch center, sending periodic updates to the
  Dorado using a setup shown in Fig. \ref{fig:dta-setup}. The Google
  Earth post-hoc image shows red track lines of the \rx commanded
  Dorado with the yellow transects of the LRAUV. Yellow dots show the
  track line of a simple GPS-enabled drifter used for seeding the
  experiment over the Monterey Canyon with Santa Cruz, CA to the right
  of the figure.}
\label{fig:ballet}
\end{figure}

Successive lagrangian experiments also followed a more innovative and
robotically choreographed ``ballet'' between two AUVs, the Dorado and
the LRAUV \cite{bellingham10} in October 2010 as part of another
series of \texttt{CANON} field experiments. The slower vehicle with a
fluorometer tracked a targeted chlorophyll patch seeded by a
GPS-enabled drifter, sending periodic updates of the patch centroid to
shore via the ODSS following a setup similar to
Fig.\ref{fig:dta-setup}, to the Dorado running \rx.

\begin{figure*}
\centering 
\subfloat[Temperature]{\label{fig:sep11-temp}\includegraphics[width=0.5\textwidth]{figs/sep11-ww-temp.png}} 
\subfloat[Salinity]{\label{fig:sep11-sal}\includegraphics[width=0.5\textwidth]{figs/sep11-ww-salinity.png}}\\
\subfloat[Fluorescence line height]{\label{fig:sep11-flh}\includegraphics[width=0.5\textwidth]{figs/sep11-ww-flh.png}} 
\subfloat[Nitrate]{\label{fig:sep11-n2}\includegraphics[width=0.5\textwidth]{figs/sep11-ww-n2.png}} 
\caption{\small Interpolated results of a September 2011 \rx mission
  in the Monterey Bay with the profiling \texttt{Wirewalker}
  \cite{rainville01,pinkel11} instrument. Plots show the instrument
  was entrained between two fronts. \emph{Image courtesy: Rishi
    Graham, MBARI}}
  \label{fig:sept11-wirewalker}
\end{figure*}

More interesting experiments are being carried out as a basis to track
more dynamic features such as ocean fronts \cite{fronts11} and
tracking more capable drifting platforms. In particular experiments
with a profiling CTD instrument, the \texttt{Wirewalker}
\cite{rainville01,pinkel11} are resulting in new modalities of making
Lagrangian and Eulerian observations. Fig. \ref{sep11-wirewalker}
shows one such experiment in September 2011 with a GPS-enabled
\texttt{Wirewalker} was tracked by the Dorado running \rx. We are
currently investigating spatio-temporal correlations of CTD data
obtained by the \texttt{Wirewalker} and tying it to the observations
made in the contextual environment by the Dorado's sensors.
